Economic Evaluation of Mis-speficied Geographically Weighted Regression Models for Site-specific Nitrogen Management

### 2023-07-27 Writing/manuscript_gwr.rmd
citation problem fixed

In recent years, on-farm precision experimentation (OFPE) has gained traction with agricultural researchers and practitioners as a means of generating high-resolution spatial data to inform site-specific crop input management [@bullock2019data; @lacoste2021farm]. Accompanying interest is surging among researchers in identifying efficient statistical methods to estimate site-specific yield response functions with OFPE data.  Researchers recently have examined various methods, ranging from Geographically Weighted Regression (GWR) [@rakshit2020; @evans2020assessment; @trevisan2021spatial] to machine learning techniques, including random forest [@krause2020random, @de2023predicting], convolutional neural network [@barbosa2020modeling], and causal forest [@kakimoto2022causal] approaches.  Among these, reported research on the GWR approach have generated special interest because GWR can be used to estimate site-specific yield-to-input relationships without costly spatial field and soil data.

@@

There are two potentially dangerous factors that may have caused significant over-estimation of the value of GWR as a tool to estimate site-specific economically optimal nitrogen application rates (EONRs) in @trevisan2021spatial. First, @trevisan2021spatial assumed a quadratic functional form in their GWR estimation even though there is a wide consensus among the scientific community that corn yield response to nitrogen is better represented by the quadratic-plateau functional form. It is likely that their choice of quadratic functional form was merely a compromise because no statistical software package implements GWR with a quadratic-plateau functional form. The second source of over-estimation bias is endemic to any study that evaluates the value of variable rate application. In general, researchers take the following steps to estimate the value of VRA using OFPE data [@anselin2004spatial, @bullock2002adding, @liu2006site, @swinton1998evaluating]: (1) estimate site-specific yield response functions based on a method chosen by the researcher; (2) use the estimated yield response functions to estimate the economically optimal site-specific input rates (EOIRs) and economically optimal uniform input rates; (3) plug the estimated site-specific EOIRs and uniform EOIR into the estimated yield response functions to calculate estimated profits under both strategies; and (4) find the difference in the two strategies' estimated profits. However, this approach can result in significant bias in the estimation of the value of VRA because of the step 3's assumption that the estimated yield response functions are the true yield response functions. @trevisan2021spatial provided an example at step (3) by assuming that their GWR estimates provided "true" site-specific yield response functions. This assumption is generally unavoidable for researchers since they cannot observe any real field's true site-specific yield response functions. But inaccurate estimation of yield response function in this methodology can result in significant bias in the estimation of the value of VRA. This bias problem can be particularly severe when the method used tends to exaggerate the true heterogeneity in yield response functions. We suspect that the high value of VRA estimated by @trevisan2021spatial displays this bias., such as GWR [@czarnota2015evaluating].

@@

An extensive literature examines the choice of functional form in yield response estimation [@cerrato1990comparison, @frank1990comparison, @bullock1994quadratic, @tembo2008crop]. Our simulation assumes that each cell $(j, i)$'s "true" yield response function  follows the commonly-used quadratic-plateau functional form:

@@

The spatial distributions of cell-level parameters $\alpha_{j, i}$, $\beta_{j, i}$, $\gamma_{j,i}$, $\tau_{j, i}$, and the error term $\varepsilon_{j, i}$ were simulated by generating spatially correlated random fields using unconditional Gaussian geostatistical simulation based on the specified variograms [@dietrich1996fast, @wood1994simulation]. The simulation processes were conducted using the R package gstat [@graler16]. Specifically, the model is spherical, the nugget variance is zero, and the range is 600 meters. The nugget value of zero means that infinitesimally close observations have a zero covariance (i.e., they are identical). The range of 600m means that observations separated by more than 600m are spatially uncorrelated. A higher range value indicates increased spatial dependence, meaning that less abrupt differences tend to be observed between the production function parameters of neighboring cells. The mean and sill variance of each parameter were specified to result in a realistic mean yield of $11,416$ kg ha$^{-1}$, and such that the cells' true economically optimal N rates ranged from $113$ kg ha$^{-1}$ to $273$ kg ha$^{-1}$ across all the simulations, a range into which most real-world N rate recommendations fall [e.g., @sawyer2018]. Note that the range of the optimal N rates of individual rounds of simulations are narrower. In defining the magnitude of the error term, we utilized the experimental data generated through eight on-farm randomized N field trials conducted by the Data-Intensive Farm Management Project (DIFM) project in 2018 [@bullock2019data]. Using the data from each of the eight actual trials, separate regressions of yield on N were run, and standard deviations of the residuals were calculated. The mean value of the eight standard deviations was $1370$ kg ha$^{-1}$, which was the square root value of the sill variances assigned to the simulations' error terms.

@@

Previous studies comparing models have predominantly focused on measuring model performance based on statistical accuracy of parameter estimates, such as bias, mean squared error, type I error, and confidence intervals, for both GWR models [@lu2017geographically; @harris2019simulation] and general econometric models [@christopeit2006local, @su2016identifying]. While these statistical measures are fundamental for evaluating model performance, for practical reasons it is imperative to consider the translation of statistical accuracy into economic value. Moreover, the extent of this translation could vary depending on the specific application scenarios. The research reported in the present article distinguishes itself from previous literature by explicitly examining the economic value of specific econometric models (GWR) under a model functional form misspecification (quadratic instead of quadratic-plateau), in a specific application (corn site-specific nitrogen management), and under specific context scenarios (normal-sized production field, various market prices, etc.), through a simulation-based methodology. We encourage other model comparison studies in empirical economic research to also incorporate similar assessment of economic value.

>>>>

In recent years, on-farm precision experimentation (OFPE) has gained traction with agricultural researchers and practitioners as a means of generating high-resolution spatial data to inform site-specific crop input management [@bullock2019data; @lacoste2021farm]. Accompanying interest is surging among researchers in identifying efficient statistical methods to estimate site-specific yield response functions with OFPE data.  Researchers recently have examined various methods, ranging from Geographically Weighted Regression (GWR) [@rakshit2020; @evans2020assessment; @trevisan2021spatial] to machine learning techniques, including random forest [@krause2020random; @de2023predicting], convolutional neural network [@barbosa2020modeling], and causal forest [@kakimoto2022causal] approaches.  Among these, reported research on the GWR approach have generated special interest because GWR can be used to estimate site-specific yield-to-input relationships without costly spatial field and soil data.

@@

There are two potentially dangerous factors that may have caused significant over-estimation of the value of GWR as a tool to estimate site-specific economically optimal nitrogen application rates (EONRs) in @trevisan2021spatial. First, @trevisan2021spatial assumed a quadratic functional form in their GWR estimation even though there is a wide consensus among the scientific community that corn yield response to nitrogen is better represented by the quadratic-plateau functional form. It is likely that their choice of quadratic functional form was merely a compromise because no statistical software package implements GWR with a quadratic-plateau functional form. The second source of over-estimation bias is endemic to any study that evaluates the value of variable rate application. In general, researchers take the following steps to estimate the value of VRA using OFPE data [@anselin2004spatial; @bullock2002adding; @liu2006site; @swinton1998evaluating]: (1) estimate site-specific yield response functions based on a method chosen by the researcher; (2) use the estimated yield response functions to estimate the economically optimal site-specific input rates (EOIRs) and economically optimal uniform input rates; (3) plug the estimated site-specific EOIRs and uniform EOIR into the estimated yield response functions to calculate estimated profits under both strategies; and (4) find the difference in the two strategies' estimated profits. However, this approach can result in significant bias in the estimation of the value of VRA because of the step 3's assumption that the estimated yield response functions are the true yield response functions. @trevisan2021spatial provided an example at step (3) by assuming that their GWR estimates provided "true" site-specific yield response functions. This assumption is generally unavoidable for researchers since they cannot observe any real field's true site-specific yield response functions. But inaccurate estimation of yield response function in this methodology can result in significant bias in the estimation of the value of VRA. This bias problem can be particularly severe when the method used tends to exaggerate the true heterogeneity in yield response functions. We suspect that the high value of VRA estimated by @trevisan2021spatial displays this bias., such as GWR [@czarnota2015evaluating].

@@

An extensive literature examines the choice of functional form in yield response estimation [@cerrato1990comparison; @frank1990comparison; @bullock1994quadratic; @tembo2008crop]. Our simulation assumes that each cell $(j, i)$'s "true" yield response function follows the quadratic-plateau functional form because our main objective is to find the bias associated with using a quadratic funcitonal form in GWR when the true yield response fucntion follows the quadratic-plateau functional form:

@@

The spatial distributions of cell-level parameters $\alpha_{j, i}$, $\beta_{j, i}$, $\gamma_{j,i}$, $\tau_{j, i}$, and the error term $\varepsilon_{j, i}$ were simulated by generating spatially correlated random fields using unconditional Gaussian geostatistical simulation based on the specified variograms [@dietrich1996fast; @wood1994simulation]. The simulation processes were conducted using the R package gstat [@graler16]. Specifically, the model is spherical, the nugget variance is zero, and the range is 600 meters. The nugget value of zero means that infinitesimally close observations have a zero covariance (i.e., they are identical). The range of 600m means that observations separated by more than 600m are spatially uncorrelated. A higher range value indicates increased spatial dependence, meaning that less abrupt differences tend to be observed between the production function parameters of neighboring cells. The mean and sill variance of each parameter were specified to result in a realistic mean yield of $11,416$ kg ha$^{-1}$, and such that the cells' true economically optimal N rates ranged from $113$ kg ha$^{-1}$ to $273$ kg ha$^{-1}$ across all the simulations, a range into which most real-world N rate recommendations fall [e.g., @sawyer2018]. Note that the range of the optimal N rates of individual rounds of simulations are narrower. In defining the magnitude of the error term, we utilized the experimental data generated through eight on-farm randomized N field trials conducted by the Data-Intensive Farm Management Project (DIFM) project in 2018 [@bullock2019data]. Using the data from each of the eight actual trials, separate regressions of yield on N were run, and standard deviations of the residuals were calculated. The mean value of the eight standard deviations was $1370$ kg ha$^{-1}$, which was the square root value of the sill variances assigned to the simulations' error terms.

@@

Previous studies comparing models have predominantly focused on measuring model performance based on statistical accuracy of parameter estimates, such as bias, mean squared error, type I error, and confidence intervals, for both GWR models [@lu2017geographically; @harris2019simulation] and general econometric models [@christopeit2006local; @su2016identifying]. While these statistical measures are fundamental for evaluating model performance, for practical reasons it is imperative to consider the translation of statistical accuracy into economic value. Moreover, the extent of this translation could vary depending on the specific application scenarios. The research reported in the present article distinguishes itself from previous literature by explicitly examining the economic value of specific econometric models (GWR) under a model functional form misspecification (quadratic instead of quadratic-plateau), in a specific application (corn site-specific nitrogen management), and under specific context scenarios (normal-sized production field, various market prices, etc.), through a simulation-based methodology. We encourage other model comparison studies in empirical economic research to also incorporate similar assessment of economic value.
